Ascension Island Implementation

### 2025-01-13 Implementation doc/Draft1.Rmd
Updating draft

source("@_Region file.R")
Period <- c("2010-2019", "2020-2029", "2030-2039", "2040-2049", "2050-2059", "2060-2069")

@@

models <- c("SSP1-CNRM", "SSP3-CNRM", "SSP1-GFDL", "SSP3-GFDL")

model <- StrathE2E2::e2e_read("SSP1_CNRM", "2010-2019", models.path = "Files/Ascension")

Area <- read.csv(str_glue("Files/Ascension/SSP1_CNRM/2010-2019/Param/event_timing_ASCENSION.csv")) %>% 
  filter(str_detect(Description, "km2")) %>% 
  .$Value

@@

physical <- read.csv(str_glue("Files/Ascension/SSP1_CNRM/2010-2019/Param/physical_parameters_ASCENSION.csv"))

@@

Using these results, we derived the necessary parameters for lines 17–26 in the event_timing_ASCENSION_*.csv files, for each decade and the four climate projections.

@@

# Ecological drivers {-}

@@

- Temperature, vertical mixing coefficients, volume fluxes, and boundary nutrient, detritus and phytoplankton concentrations from outputs of an ensemble of NEMO-ERSEM coupled hydro-geochemical model runs with a 2015 historical/future split CITE YURI XXX.
- Surface shortwave radiation from CMIP6 used to force the NEMO-ERSEM coupled hydro-geochemical runs mentioned above.
- Freshwater volume outflows from CMIP6 used to force the NEMO-ERSEM coupled hydro-geochemical runs mentioned above.
- River nitrate and ammonia concentrations taken from the IMAGE model to force the NEMO-ERSEM coupled hydro-geochemical runs mentioned above.
- Atmospheric deposition of nitrate and ammonia from ISIMIP3a ([Inter-Sectoral Impact Model Intercomparison Project](https://data.isimip.org/10.48364/ISIMIP.759077.2); [@yang22]) with SSP projects from ISIMIP3b used to force the NEMO-ERSEM coupled hydro-geochemical runs mentioned above.
- Remote sensing data products on Suspended Particulate Matter (Globcolour L3b; ftp://ftp.hermes.acri.fr/GLOB/merged/month/).
- Wave height, period, and direction from the ERA-5 reanalysis monthly means accessed through [CDS](https://cds.climate.copernicus.eu/cdsapp#!/dataset/10.24381/cds.f17050d7?tab=overview) for 1980-2021 [@Hersbach19].

@@

### NEMO-ERSEM ensemble: ### {-}

@@

### Vertical mixing coefficients between the model layers: ### {-}

@@

### Monthly averaged temperatures for each water column layer: ### {-}

@@

### Monthly averaged suspended particulate matter (SPM) concentrations (mg.m^-3^) in the shallow zone and the deep zone upper layer: ### {-}

Monthly averaged values of inorganic suspended particulate matter in sea water are available from the Globcolour project, starting from September 1997. These data are derived from satellite observations using the algorithm of Gohin [-@gohin11]. Data were downloaded from the ftp server (ftp://ftp.hermes.acri.fr/GLOB/merged/month/). We summarised these values as zonal statistics for the model domain to acquire a climatological annual cycle of data for the 2010-2019 simulation period only.

### Monthly average light attenuation coefficients for the inshore and offshore surface layers: ### {-}

@@

### Monthly averaged daily integrated irradiance at the sea surface (E.m^-2^.d^-1^): ### {-}

@@

### Monthly averaged daily atmospheric deposition rates of oxidised and reduced nitrogen onto the sea surface in the shallow and deep zones (mMN.m^-2^.d^-1^): ### {-}

@@

### Monthly averaged, freshwater river inflow rates (expressed as a daily proportion of the receiving layer volume), and concentrations of oxidised and reduced dissolved inorganic nitrogen in the inflowing river waters (mMN.m^-3^): ### {-}

@@

### Volume fluxes into the model domain across open sea boundaries, and from the upper layer of the offshore/deep zone into the inshore/shallow zone, expressed as proportions of the receiving layer volume per day: ### {-}

@@

### Mean concentrations of nitrate, ammonia, phytoplankton and suspended detritus (mMN.m^-3^), in adjacent ocean waters inflowing to the offshore/deep zone upper layer, adjacent ocean waters inflowing to the offshore/deep zone lower layer, and adjacent shelf waters inflowing to the inshore/shallow zone: ### {-}

@@

Ferrari et al. (2024a) noted a high biomass of fish within 31–60 cm and above 60 cm, which could be linked to some level of protection from fishing and favorable environmental conditions. Among large predators, Ascension Island supports high densities of the rock hind/grouper (Choat and Robertson, 2008). Surveys conducted by Choat and Robertson (2006) in sub-littoral zones estimated an average density of 0.078 individuals ± 0.005 per m² in 2006. A subsequent survey by the Marine and Fisheries Unit of the Ascension Island Government in 2015 recorded a higher density of 0.205 individuals ± 0.02 per m² (Ascension Island Government, 2016). While these figures could suggest an increase in grouper abundance, the authors cautioned that the surveys took place during a fry event, which may have influenced the results (Ascension Island Government, 2016). Although biomass estimates for the grouper are not yet available, we used the reported abundance figures to approximate its contribution to the total biomass (average the values from Quimbayo et al., 2018 and Ferrari et al., 2024). By applying species-specific length data from FishBase and using Bayesian length-weight equations (W = a * LT^b, where a and b are constants specific to Ascension Island), we estimate that groupers could account for over 12% of the total biomass when considering abundance values from Choat and Robertson (2006). When using the data from the Ascension Island Government (2016), this contribution rises to 32.6%.

To establish a target production value for demersal fish in the model we calculated a rough biomass estimate of 76,010.5 tonnes based on the average of the reef fish biomass in the studies of Quimbayo et al. (2018) and Ferrari et al. (2024). To calculate total biomass in the inshore area, we used an area of 108.2 km² (reef area <200 m depth sourced from Ferrari et al., 2023). We applied a PB value of 0.76, based on the M. niger PB value available in Capitani et al. (2021).
This yielded a production value of 57,767.98 tons.

@@

## Diet proportions {-}

@@

### Seabirds {-}

>>>>

source("@_Region file.R")

@@

models <- StrathE2E2::e2e_ls("Files/")

model <- StrathE2E2::e2e_read("Ascension_MA", "2010-2019-CNRM-ssp126", models.path = "Files")

Area <- read.csv(str_glue("Files/Ascension_MA/2010-2019-CNRM-ssp126/Param/event_timing_ASCENSION_MA_2010-2019-CNRM-ssp126.csv")) %>% 
  filter(str_detect(Description, "km2")) %>% 
  .$Value

@@

physical <- read.csv(str_glue("Files/Ascension_MA/2010-2019-CNRM-ssp126/Param/physical_parameters_ASCENSION.csv"))

@@

Using these results, we derived the necessary parameters for lines 17–26 in the event_timing_ASCENSION\_\*.csv files, for each decade and the four climate projections.

@@

# Ecological drivers {.unnumbered}

@@

-   Temperature, vertical mixing coefficients, volume fluxes, and boundary nutrient, detritus and phytoplankton concentrations from outputs of an ensemble of NEMO-ERSEM coupled hydro-geochemical model runs with a 2015 historical/future split CITE YURI XXX.
-   Surface shortwave radiation from CMIP6 used to force the NEMO-ERSEM coupled hydro-geochemical runs mentioned above.
-   Freshwater volume outflows from CMIP6 used to force the NEMO-ERSEM coupled hydro-geochemical runs mentioned above.
-   River nitrate and ammonia concentrations taken from the IMAGE model to force the NEMO-ERSEM coupled hydro-geochemical runs mentioned above.
-   Atmospheric deposition of nitrate and ammonia from ISIMIP3a ([Inter-Sectoral Impact Model Intercomparison Project](https://data.isimip.org/10.48364/ISIMIP.759077.2); [@yang22]) with SSP projects from ISIMIP3b used to force the NEMO-ERSEM coupled hydro-geochemical runs mentioned above.
-   Remote sensing data products on Suspended Particulate Matter (Globcolour L3b; <ftp://ftp.hermes.acri.fr/GLOB/merged/month/>).
-   Wave height, period, and direction from the ERA-5 reanalysis monthly means accessed through [CDS](https://cds.climate.copernicus.eu/cdsapp#!/dataset/10.24381/cds.f17050d7?tab=overview) for 1980-2021 [@Hersbach19].

@@

### NEMO-ERSEM ensemble: {.unnumbered}

@@

### Vertical mixing coefficients between the model layers: {.unnumbered}

@@

### Monthly averaged temperatures for each water column layer: {.unnumbered}

@@

### Monthly averaged suspended particulate matter (SPM) concentrations (mg.m^-3^) in the shallow zone and the deep zone upper layer: {.unnumbered}

Monthly averaged values of inorganic suspended particulate matter in sea water are available from the Globcolour project, starting from September 1997.
These data are derived from satellite observations using the algorithm of Gohin [-@gohin11].
Data were downloaded from the ftp server (<ftp://ftp.hermes.acri.fr/GLOB/merged/month/>).
We summarised these values as zonal statistics for the model domain to acquire a climatological annual cycle of data for the 2010-2019 simulation period only.

### Monthly average light attenuation coefficients for the inshore and offshore surface layers: {.unnumbered}

@@

### Monthly averaged daily integrated irradiance at the sea surface (E.m^-2^.d^-1^): {.unnumbered}

@@

### Monthly averaged daily atmospheric deposition rates of oxidised and reduced nitrogen onto the sea surface in the shallow and deep zones (mMN.m^-2^.d^-1^): {.unnumbered}

@@

### Monthly averaged, freshwater river inflow rates (expressed as a daily proportion of the receiving layer volume), and concentrations of oxidised and reduced dissolved inorganic nitrogen in the inflowing river waters (mMN.m^-3^): {.unnumbered}

@@

### Volume fluxes into the model domain across open sea boundaries, and from the upper layer of the offshore/deep zone into the inshore/shallow zone, expressed as proportions of the receiving layer volume per day: {.unnumbered}

@@

### Mean concentrations of nitrate, ammonia, phytoplankton and suspended detritus (mMN.m^-3^), in adjacent ocean waters inflowing to the offshore/deep zone upper layer, adjacent ocean waters inflowing to the offshore/deep zone lower layer, and adjacent shelf waters inflowing to the inshore/shallow zone: {.unnumbered}

@@

Ferrari et al. (2024a) noted a high biomass of fish within 31–60 cm and above 60 cm, which could be linked to some level of protection from fishing and favorable environmental conditions.
Among large predators, Ascension Island supports high densities of the rock hind/grouper (Choat and Robertson, 2008).
Surveys conducted by Choat and Robertson (2006) in sub-littoral zones estimated an average density of 0.078 individuals ± 0.005 per m² in 2006.
A subsequent survey by the Marine and Fisheries Unit of the Ascension Island Government in 2015 recorded a higher density of 0.205 individuals ± 0.02 per m² (Ascension Island Government, 2016).
While these figures could suggest an increase in grouper abundance, the authors cautioned that the surveys took place during a fry event, which may have influenced the results (Ascension Island Government, 2016).
Although biomass estimates for the grouper are not yet available, we used the reported abundance figures to approximate its contribution to the total biomass (average the values from Quimbayo et al., 2018 and Ferrari et al., 2024).
By applying species-specific length data from FishBase and using Bayesian length-weight equations (W = a \* LT\^b, where a and b are constants specific to Ascension Island), we estimate that groupers could account for over 12% of the total biomass when considering abundance values from Choat and Robertson (2006).
When using the data from the Ascension Island Government (2016), this contribution rises to 32.6%.

To establish a target production value for demersal fish in the model we calculated a rough biomass estimate of 76,010.5 tonnes based on the average of the reef fish biomass in the studies of Quimbayo et al. (2018) and Ferrari et al. (2024).
To calculate total biomass in the inshore area, we used an area of 108.2 km² (reef area \<200 m depth sourced from Ferrari et al., 2023).
We applied a PB value of 0.76, based on the M.
niger PB value available in Capitani et al. (2021).
This yielded a production value of 57,767.98 tons.

@@

## Diet proportions {.unnumbered}

@@
